How to write a PhD in Biological Sciences: a guide for the uninitiated

### 2021-07-20 2.09-databases.Rmd
Add files via upload
# After copy edits

Databases of scientific literature have been around since the 1960s when the number of journals and papers started increasing so much that scholars simply couldn’t keep up. Quite simply, you should use one or more of these databases to look for relevant literature to read. The first database, and what continues to be one of the largest, was the Science Citation Index (now renamed [Web of Science](https://www.webofknowledge.com/)) that sought to record not only the contents of selected journals (titles, abstracts, authors and other metadata), but also the citations that each paper used. As we have seen in the last chapter, citations record the way in which people find the literature useful, and are therefore a highly effective way of finding related and pertinent literature in your field.

Before you start using a literature database, you will need to have some key literature with which you are familiar and that relates directly to your project. If you don’t have this, then check with your advisor. Often, this key literature is cited in the advert for the PhD, the grant application that funded it, or if this is a topic that your advisor has already conducted some research on, then their own publications may include part of the key literature. If you are not familiar with your advisor’s (including co-advisor’s where appropriate) papers, then a great first step is to search for them on a database and read through titles and abstracts that are relevant to your research topic. Once you have a set of key literature (this might only be 5 to 10 papers), then it is time to go to the literature databases to do your search.

@@

## Keywords

While searching for keywords sounds easy, picking the keywords that you should use is more difficult and requires a lot of practice with the databases and your key references. Remember that keywords are often not single words, but also include short phrases. Look at all of the key references and the keywords that they contain. Any keywords that they share are going to be important for you. Make a note of all relevant keywords for your searches, as ultimately this will help you when you need to determine the keywords for [your own chapters](title.html).

@@

- Use AND when you want both terms included. For example, *performance AND lizards*. In this example, all papers that mention performance AND lizards in the title, abstract or keywords are returned. 
- Use OR when you want to include either term. For example, *performance AND lizards OR geckos*. This example should return more literature than the previous search as it will include, all papers that mention *performance AND lizards OR geckos* in the title, abstract or keywords are returned. You will note that geckos are lizards, but people who write about geckos might not mention that they are lizards in the title, abstract or keywords, so you may need to use these extra search terms in order to pick up all of the relevant literature on lizards. If your research requires you to cover all lizards, then you might need a lot of search terms in the OR string in order to pick up all of the literature that covers lizard performance.
- Use NOT when you specifically want to exclude a keyword. Following our example above, the search term *performance AND reptiles NOT snakes* should give you all performance papers on lizards, but you may want to exclude turtles and crocodilians as well if they have a lot of hits.

@@

You can also use the term NEAR to indicate that two terms need to occur within a specified number of words. For example, performance NEAR/5 reptile will capture items where these terms occur within 5 words of each other, but no greater. This might be useful to exclude some papers that mention performance and reptiles in the abstract, but are not about reptile performance.

Once you start putting together search strings you will need to use brackets to make sure that the correct terms are grouped. For example, performance AND (lizards OR geckos) will get what we described above, whereas (performance AND lizards) OR geckos will give you all papers on lizards and performance plus all papers on geckos (not what we wanted).

@@

- Use * when you want to replace a set of letters at the end of a word. For example, _reptil*_ will return all terms including Reptilia, reptile and reptiles.
- Use ? when you want to search for any single letter at that place in a word. For example, _sterili?e_ will capture words with sterilize and sterilise.
- Use $ when there is more than a single letter that might change within a word, or the omission of a letter. For example, _colo$r_ will return both US and UK spelling, color and colour, respectively.

@@

An invaluable way in which databases have become extremely useful is the ability to move items that you have found during your search directly into your [reference manager](referencesoftware.html). Beware that these regularly come with errors, and that the amount of errors is related to your database source (Scopus and WoS are much better than Google Scholar). This means that after importing the refrences, you will need to go back to your reference manager and edit them or they will come out incorrectly when you use them. Of particular note to biologists is the need to have all species names in [italics](taxonomy.html), and you'll need to do this yourself.

@@

In addition to using keywords, you should search citations of your key references. Typically, when you start reviewing the literature in your specific area you will have a few key references that have been given to you by your advisor. Having read these, you will already know that they cite other works of interest, and you should follow up to look at all that seem relevant. But that’s not all, these key publications are likely to have also been cited themselves by newer papers. Tracking citations of key papers can help you keep abreast of what is happening in your field. Indeed, given the prolific ever increasing nature of scientific publications, it might be the only way to keep on top of what is happening. Some literature databases allow you to receive alerts if a paper is cited, and you should set this up for all of your key citations, and perhaps for some of the most prominent authors in your field.

@@

The above section is a brief guide, and your library is likely to have in depth guides on how to use the specific databases that they house. They may also provide training sessions, or online videos to watch. It is worth becoming familiar with exactly how the databases that you will use work, so this is time well spent at the beginning of your studies.

@@

Citations provide an idea of how the literature is used. This can be used to assess not simply the productivity of a scientist (i.e. number of publications), but their usefulness by others (i.e. number of citations).

Clearly, the Science Citation Index cost some money to put together, and universities paid annually to receive the books, and then the CDs and finally to access the online versions. They were not for profit organisations, but eventually they were bought out by the evil profit making parasites of science. Yes, publishers saw the dependency that scientists have on a resource and decided to own it for themselves and then radically increase the price to access it.

@@

In the life sciences, there are three prominent databases used: Google Scholar, Web of Science (owned by Clarivate Analytics aka Thomson Reuters) and Scopus (owned by Elsevier). But new players are entering the market all the time, and two of the newest are Dimensions from Digital Science and Microsoft Academic.

@@

Once you have found a scientist who works within your field and produces influential work that you are keen to read, then you may want to search for all of their other papers to determine whether there is more literature that you have missed. The databases make such searches easy, but you will need the (correct) spelling of your scientist's name, and also to be aware that they may have used different iterations of their name throughout their career.

@@

While Dave Thompson could always add his middle initial to specify his work, as time went by there were more and more scientists publishing as D J Thompson. This made it more difficult for administrators and scientits alike. This has begun to be a bigger global problem, such that the databases have begun to use profiles of particular scientists that they can 'own' and curate. Each database has its own way of doing this, and it can become rather a chore so that not many scientists have their own profiles, which makes it a problem particularly for those with common names.

The [ORCID](orcid.html) number has been errected to provide a common platform for [authors to curate](title.html), and now some journals and funding authorities won't allow you to submit or apply without one.

@@

Well it seems that there are some people who have already thought about this, and come up with a good idea of exactly what's different. Martín-Martín et al [-@martin-martin2018google] have done a great job of analysing all this stuff from some 2.5 million citations. What they found inspired me to write this chapter, in which I've focussed on the results from the Life-Sciences. But I encourage you to go read the article for yourself.

I have been known to take the odd peak at my Google Scholar profile over the year, and see how the citations are accumulating. I rarely check on WoS or Scopus, because it's a bit of a faff getting signed in and doing the search. Plus it looks so much smaller when one is habituated to seeing those double digits in GS! However, I've always been a bit uneasy about citing my GS citation rate, H-index or i10 (among others that they give) as I've never really known what all that extra represents. Something grey and unseemly? Well, it turns out that it's all good stuff, and perhaps GS is the better one to cite as it's a more inclusive index: more inclusive of different document types and different languages.

@@

To directly compare your citation metrics from different platforms, and also to get more unbiased metrics than these platforms provide, consider using the software Publish or Perish [@harzing2007publish]. This also provides a great platform on which to harvest data from Google Scholar into a database, or put it directly into another database, like EndNote, Zotero or Mendelay.

@@

According to Martín-Martín et al [-@martin-martin2018google] The GS score is likely to be inflated - 1.90 for GS/WoS and 1.45 for GS/Scopus. If you deviate from this with a higher score, you can give yourself a pat on the back for having work that's reaching more people in more parts of the world. Google Scholar results reflect a more inclusive citation index [@martin-martin2018google]. While WoS and Scopus aren't exclusively English or journal publications, they are mostly. But that extra third that GS gives you allows you to show the extra scope that your work is getting outside that English journal mainstream. Is your GS score more than a third higher than your WoS or Scopus score? If yes, then your work is having a greater impact elsewhere in the world, and there's nothing wrong with that.

>>>>

Databases of scientific literature have been around since the 1960s, when the number of journals and papers started increasing so much that scholars simply couldn’t keep up. Quite simply, you should use one or more of these databases to look for relevant literature to read. The first database, and one that continues to be one of the largest, was the Science Citation Index (now renamed [Web of Science](https://www.webofknowledge.com/)) that sought to record not only the contents of selected journals (titles, abstracts, authors and other metadata), but also the citations that each paper used. As we have seen in the last chapter, citations record the way in which people find the literature useful, and are therefore, a highly effective way of finding related and pertinent literature in your field.

Before you start using a literature database, you will need to have some key literature with which you are familiar and that relates directly to your project. If you don’t have this, then check with your advisor. Sometimes, this key literature is cited in the advert for the PhD, the grant application that funded it, or if this is a topic that your advisor has already conducted some research on, then their own publications may include part of the key literature. If you are not familiar with your advisor’s (including co-advisor’s where appropriate) papers, then a great first step is to search for them on a database and read through titles and abstracts that are relevant to your research topic. Once you have a set of key literature (this might only be 5 to 10 papers), then it is time to go to the literature databases to do your search.

@@

## Keywords {#keywords}

While searching for keywords sounds easy, picking the keywords that you should use is more difficult and requires a lot of practice with the databases and your key references. Remember that keywords are often not single words, but also include short phrases. Look at all of the key references and the keywords that they contain. Any keywords that they share are going to be important for you. Make a note of all relevant keywords for your searches, as ultimately, this will help you when you need to determine the keywords for [your own chapters](title.html).

@@

- Use AND when you want both terms included. For example, *performance AND lizards*. In this example, all papers that mention performance AND lizards in the title, abstract or keywords are returned. 
- Use OR when you want to include either term. For example, *performance AND lizards OR geckos*. This example should return more literature than the previous search as it will include, all papers that mention *performance AND lizards OR geckos* in the title, abstract or keywords are returned. You will note that geckos are lizards, but people who write about geckos might not mention that they are lizards in the title, abstract or keywords, so you may need to use these extra search terms in order to pick up all of the relevant literature on lizards. If your research requires you to cover all lizards, then you might need a lot of search terms in the OR string in order to pick up all of the literature that covers lizard performance.
- Use NOT when you specifically want to exclude a keyword. Following our example above, the search term *performance AND reptiles NOT snakes* should give you all performance papers on lizards, but you may want to exclude turtles and crocodilians as well if they have a lot of hits. *NOT* is a powerful operator and needs to be used with caution, as it could exclude some relevant hits. In our example here, an item about interactions between snakes and lizards would be excluded.

@@

You can also use the term NEAR to indicate that two terms need to occur within a specified number of words. For example, *performance* NEAR/5 *reptile* will capture items where these terms occur within 5 words of each other, but no greater. This might be useful to exclude some papers that mention performance and reptiles in the abstract but are not about reptile performance.

Once you start putting together search strings you will need to use brackets to make sure that the correct terms are grouped. For example, *performance* AND (*lizards* OR *geckos*) will get what we described above, whereas (*performance* AND *lizards*) OR *geckos* will give you all papers on lizards and performance plus all papers on geckos (not what we wanted).

@@

- Use * when you want to replace a set of letters at the end of a word. For example, _reptil*_ will return all terms including Reptilia, reptile and reptiles.
- Use ? when you want to search for any single letter at that place in a word. For example, _sterili?e_ will capture words with sterilize and sterilise.
- Use $ when there is more than a single letter that might change within a word, or the omission of a letter. For example, _colo$r_ will return both US and UK spelling, color and colour, respectively. Note that Scopus doesn't use the $ symbol as a wildcard: for this example you can use ? instead.

@@

An invaluable way in which databases have become extremely useful is the ability to move items that you have found during your search directly into your [reference manager](referencesoftware.html). Beware that these regularly come with errors, and that the amount of errors is related to your database source (Scopus and WoS are much better than Google Scholar). This means that after importing the references, you will need to go back to your reference manager and edit them or they will come out incorrectly when you use them. Of particular note to biologists is the need to have all species names in [italics](taxonomy.html), and you'll need to do this yourself.

@@

In addition to using [keywords](#keywords), you should search citations of your key references. Typically, when you start reviewing the literature in your specific area, you will have a few key references that have been given to you by your advisor. Having read these, you will already know that they cite other works of interest, and you should follow up to look at all that seem relevant. But that’s not all, these key publications are likely to have also been cited themselves by newer papers. Tracking citations of key papers can help you keep abreast of what is happening in your field. Indeed, given the prolific ever increasing nature of scientific publications, it might be the only way to keep on top of what is happening. Some literature databases allow you to receive alerts if a paper is cited (e.g. Google Scholar), and you should set this up for all of your key citations, and perhaps for some of the most prominent authors in your field.

@@

The above section is a brief guide, and your library is likely to have in-depth guides on how to use the specific databases that they house. They may also provide training sessions, or online videos to watch. It is worth becoming familiar with exactly how the databases that you will use work, so this is time well spent at the beginning of your studies.

@@

Citations provide an idea of how the literature is used. This can be used to assess not simply the productivity of a scientist (i.e. number of publications), but their usefulness to others (i.e. number of citations).

Clearly, the Science Citation Index cost some money to put together, and universities paid annually to receive the books, and then the CDs and finally to access the online versions. They were a not for profit organisation, but eventually they were bought out.

@@

In the life sciences, there are three prominent databases used: Google Scholar, Web of Science and Scopus. But new players are entering the market all the time, and two of the newest are Dimensions and Microsoft Academic.

@@

Once you have found a scientist who works within your field and produces influential work that you are keen to read, then you may want to search for all of their other papers to determine whether there is more literature that you might have missed. The databases make such searches easy, but you will need the (correct) spelling of your scientist's name, and also to be aware that they may have used different iterations of their name throughout their career.

@@

While Dave Thompson could always add his middle initial to specify his work, as time went by there were more and more scientists publishing as D J Thompson. This made it more difficult for administrators and scientists alike. This has begun to be a bigger global problem, such that the databases have begun to use profiles of particular scientists that individuals can 'own' and curate. Each database has its own way of doing this, and it can become rather a chore so that not many scientists have their own profiles, which makes it a problem, particularly for those with common names.

The [ORCID](orcid.html) number has been erected to provide a common platform for [authors to curate](title.html), and now some journals and funding authorities won't allow you to submit or apply without one.

@@

Well, it seems that there are some people who have already thought about this, and come up with a good idea of exactly what's different. Martín-Martín et al. [-@martin-martin2018google] have done a great job of analysing some 2.5 million citations. What they found inspired me to write this section, in which I've focussed on the results from the Life-Sciences. But I encourage you to go read the article for yourself.

I have been known to take the odd peek at my Google Scholar profile over the year, and see how the citations are accumulating. I rarely check on WoS or Scopus, because it's a bit of a faff getting signed in and doing the search. Plus it looks so much smaller when one is habituated to seeing those double digits in GS! However, I've always been a bit uneasy about citing my GS citation rate, H-index or i10 (among other mertics that they give) as I've never really known what all that extra represents. Something grey and unseemly? Well, it turns out that it's all good stuff, and perhaps GS is the better one to cite as it's a more inclusive index: more inclusive of different document types and different languages.

@@

To directly compare your citation metrics from different platforms, and also to get more unbiased metrics than these platforms provide, consider using the software Publish or Perish [@harzing2007publish]. This also provides a great platform on which to harvest data from Google Scholar into a database, or put it directly into another database, like EndNote, Zotero or Mendeley.

@@

According to Martín-Martín et al. [-@martin-martin2018google], the GS score is likely to be inflated - 1.90 for GS/WoS and 1.45 for GS/Scopus. If you deviate from this with a higher score, you can give yourself a pat on the back for having work that's reaching more people in more parts of the world. Google Scholar results reflect a more inclusive citation index [@martin-martin2018google]. While WoS and Scopus aren't exclusively English or journal publications, they are mostly. But that extra third that GS gives you allows you to show the extra scope that your work is getting outside that English journal mainstream. Is your GS score more than a third higher than your WoS or Scopus score? If yes, then your work is having a greater impact elsewhere in the world, and there's nothing wrong with that.

### 2021-07-20 2.10-referencesortware.Rmd
Add files via upload
# After copy edits

Because published papers play such a fundamental role in your [reading and comprehension](critical.html) of a subject, and [citations](citations.html) are pivotal in chapter or paper writing, putting these elements together in a good reference manager makes a lot of sense. This is a piece of software in which you can store and view all of your papers (in pdf format or via links), make notes on them, and use them to make citations and build bibliographies. Unsurprisingly, they become something that you may well use a lot during the course of your PhD, and it is worth thinking carefully about which one you want to use.

There are quite a lot of different software packages out there to help you do this (e.g. [Endnote](https://endnote.com/), [Mendelay](https://www.mendeley.com/guides/desktop), [Zotero](https://www.zotero.org/) - see more about these [here](https://subjectguides.library.american.edu/c.php?g=479020&p=3323781)). These are a mixed blessing. Firstly, most of them are not free, so unless your university subscribes to their providers you would have to pay for the privilege of using them. Next, all of the ones that I've ever used are very demanding and do not necessarily work well with your word processor of choice. Given all the time I have invested in reference formatting software, I think I might have been better off just copying and pasting references I have at the end of a document and then formatting them by hand. This tends to be how I do it now. Having said this, I used Zotero as the referencing software for this book. It took very little time to learn how to use, it was painless to enter (most) of the references into the database, and it is free and works with [Bookdown](bookdown.org) and Google Docs.

>>>>

Because published papers play such a fundamental role in your [reading and comprehension](critical.html) of a subject, and [citations](citations.html) are pivotal in chapter or paper writing, putting these elements together in a good reference manager makes a lot of sense. A reference manager is a piece of software in which you can store and view all of your papers (in pdf format or via links), make notes on them, and use them to make citations and build bibliographies. Unsurprisingly, they become something that you may well use a lot during the course of your PhD, and it is worth thinking carefully about which one you want to use.

There are quite a lot of different software packages out there to help you do this (e.g. [Endnote](https://endnote.com/), [Mendeley](https://www.mendeley.com/guides/desktop), [Zotero](https://www.zotero.org/) - see more about these [here](https://subjectguides.library.american.edu/c.php?g=479020&p=3323781)). These packages are a mixed blessing. Firstly, most of them are not free, so unless your university subscribes to their providers you would have to pay for the privilege of using them. Next, all of the ones that I've ever used are very demanding and do not necessarily work well with your word processor of choice. Given all the time I have invested in reference formatting software, I think I might have been better off just copying and pasting references I have at the end of a document and then formatting them by hand. This tends to be how I do it now. Having said this, I used Zotero as the referencing software for this book. It took very little time to learn how to use, it was painless to enter (most) of the references into the database, and it is free and works with [Bookdown](bookdown.org) and Google Docs.

### 2021-07-20 2.10a-sortware.Rmd
Add files via upload
# After copy edits

The simple answer is that you should use whatever you, and your advisor, feel most comfortable with. You will face a lot of challenges during your PhD including learning to write with lots of different kinds of software. Each one you learn to use will require a steep learning curve, and that will require your time. If there's any good time to start with a new word processing package it will be before you start your PhD. If you do decide to go with a new writing package then my suggestion would be to choose open source software like [Rmarkdown @xie2018r]. More on these later.

Another important consideration is whether or not your chosen word processing package will allow seemless integration with your [reference manager](referencesoftware.html). Your lab may have a communal repository of references, and so this may then constrain what you are comfortable using.

If you feel comfortable with a word processing package with a nice GUI, like Microsoft Word, then go ahead and use it. These certainly have advantages like being able to get a synonym at the click of a mouse (Figure \@ref(fig:synonym)). My best suggestion is that you do not save your files in their default ".docx" format. Instead make sure that all of your files are saved as .rtf or html. This should make little difference to your opening and editing but will mean that all of your texts are forward compatible with other kinds of software. It will reduce your file size and reduce the chances of file corruption: i.e. when part of the file malfunctions and you lose the contents. That’s right, in addition to loss of data by back up deficiencies [see Part III](data.html), you also have the risk of having corrupt files that will also mean you lose all of the contents. Even if you always use Word and never use any other file format, do yourself a favour and save each version as a .rtf or .txt file as well. Just like a back up, this means that you won’t have lost all of the text if a file goes corrupt. And yes, I do know people who have multiple backup versions of corrupt files: it’s happened to me.

@@

Many academics are turning to latex in order to write their manuscripts as this gives a lot of possibility to insert links, formulae, references, images, etc. in a very simple way. Once you are ready to print a latex document it can be shared very easily as a PDF.

Be aware that the problems may come not with what you want but with how your advisor and or co-authors are comfortable with. I find that commenting on a PDF takes more time than doing the same on a Word document. I don't mind working with latex but I know some advisors are very adverse to having to use this platform. Obviously in order to get the feedback from your advisor you are going to need to use a platform that they are also comfortable with. So please consult them on what they are happy to use. You may also need to instruct them about how to make comments. For example track changes is a function that a lot of people like to use but it doesn't translate back into other platforms seamlessly.

As with many things in this book I urge you to keep an open mind with respect to software. For example if your lab is progressive and already experimenting with new and flexible software platforms then I suggest that you learn with them. The future of computing is certainly likely to be open-source software and so the sooner that you become more familiar with this, the easier you will find it to adapt in future.

@@

Importantly, what you want to try to avoid is conducting a major switch between software right at the end of your PhD. Try making such changes when you have more time, like after you’ve finished!

@@

Good file management is an important aspect of a PhD. I like my students to give each chapter of their thesis a nickname or short running title that we both agree on to name the files. An example might be “Swimming_performance_V1.rft”. The reason why I don’t like “Chapter_1.rtf” as a filename is that every student has a chapter 1, so I can’t distinguish between them. In addition, I’m not good at remembering which chapter is chapter 1 and which is chapter 2. Sometimes, towards the end of a thesis, the chapters get their orders switched around anyway, causing even more confusion. Notice that I don’t like having spaces in filenames. This is a hangover from old days of MS-DOS, but is consistently useful today when using R and other coding software. However, some software platforms react oddly with underscore (_), and so you might need to work with a hyphen instead of a space (-).

Once I’ve commented on a file and sent it back in an email, or uploaded it to their folder, I’ll add my initials Swimming_performance_V1_JM.rtf. Each time I comment on this version, I will add a number to my initials and the student adds one back, so eventually we might come to: Swimming_performance_V1_JM3_MM2.rtf. After three rounds, we would normally send a clean V2 out to co-authors and once their comments are back in we should be all set.

>>>>

The simple answer is that you should use whatever you, and your advisor, feel most comfortable with. You will face a lot of challenges during your PhD including learning to write with lots of different kinds of software. Each one you learn to use will require a steep learning curve, and that will require your time. If there's any good time to start with a new word processing package it will be before you start your PhD. If you do decide to go with a new writing package then my suggestion would be to choose open source software like [R Markdown @xie2018r]. More on these later.

Another important consideration is whether or not your chosen word processing package will allow seamless integration with your [reference manager](referencesoftware.html). Your lab may have a communal repository of references, and so this may then constrain what you are comfortable using.

If you feel comfortable with a word processing package with a nice GUI, like Microsoft Word, then go ahead and use it. These certainly have advantages like being able to get a synonym at the click of a mouse (Figure \@ref(fig:synonym)). My best advice is that you do not save your files in their default ".docx" format. Instead, make sure that all of your files are saved as .rtf, .txt or .html. This should make little difference to your opening and editing but will mean that all of your texts are forward compatible with other kinds of software. It will reduce your file size and reduce the chances of file corruption: i.e. when part of the file malfunctions and you lose the contents. That’s right, in addition to loss of data by back up deficiencies [see Part III](data.html), you also have the risk of having corrupt files that will also mean you lose all of the contents. Even if you always use Word and never use any other file format, do yourself a favour and save each version as a .rtf or .txt file as well. Just like a back up, this means that you won’t have lost all of the text if a file goes corrupt. And yes, I do know people who have multiple backup versions of corrupt files: it’s happened to me.

@@

Many academics are turning to LaTeX in order to write their manuscripts as this gives a lot of possibilities to insert links, formulae, references, images, etc. in a very simple way. Once you are ready to print a LaTeX document it can be shared very easily as a PDF.

Be aware that the problems may come not with what you want but with how your advisor and or co-authors are comfortable. I find that commenting on a PDF takes more time than doing the same on a Word document. I don't mind working with LaTeX but I know some advisors are very adverse to having to use this platform. Obviously, in order to get the feedback from your advisor you are going to need to use a platform that they are also comfortable with. So please consult them on what they are happy to use. You may also need to instruct them about how to make comments. For example, track changes is a function that a lot of people like to use but it doesn't translate back into other platforms seamlessly.

As with many things in this book I urge you to keep an open mind with respect to software. For example, if your lab is progressive and already experimenting with new and flexible software platforms then I suggest that you learn with them. The future of computing is certainly likely to be open-source software, and so the sooner that you become more familiar with this, the easier you will find it to adapt in future.

@@

Importantly, what you want to try to avoid is conducting a major switch between software right at the end of your PhD. Try making such changes when you have more time.

@@

Good file management is an important aspect of a PhD. I like my students to give each chapter of their thesis a nickname or short running title that we both agree on to name the files. An example might be “Swimming_performance_V1.rtf”. The reason why I don’t like “Chapter_1.rtf” as a filename is that every student has a chapter 1, so I can’t distinguish between them. In addition, I’m not good at remembering which chapter is chapter 1 and which is chapter 2. Sometimes, towards the end of a thesis, the chapters get their orders switched around, causing even more confusion. Notice that I don’t like having spaces in filenames. This is a hangover from old days of MS-DOS, but is consistently useful today when using R and other coding software. However, some software platforms react oddly with underscore (_), and so you might need to work with a hyphen instead of a space (-).

Once I’ve commented on a file and sent it back in an email, or uploaded it to their shared folder, I’ll add my initials Swimming_performance_V1_JM.rtf. Each time I comment on this version, I will add a number to my initials and the student adds one back, so eventually, we might come to: Swimming_performance_V1_JM3_MM2.rtf. After three rounds, we would normally send a clean V2 out to co-authors and once their comments are back in we should be all set.

### 2021-07-20 2.11-paywall.Rmd
Add files via upload
# After copy edits

Once you've understand where the money for science and publishing comes from, you will understand that there is no real shame in trying to access papers that are behind a paywall. Essentially, I’m going to recommend that you move from legitimate to illegitimate, and then if all else fails to get an interlibrary loan. This is last on the list as someone will have to pay for it. Your library may have an allocation that they can use before the bill comes to your advisor’s desk.

@@

Google scholar has lots of links (on the right) labelled with “[pdf]”, and these are often integrated with your library. If there’s nothing at first, click on the “All 9 versions” link (below and to the right). This may reveal that there is more than one version with a pdf, and you might have access to one of the others. Check for a link to the author on google scholar, and then look for a link to their website.

@@

@@

There are a few of these, and some have plugins to web browsers. You may need to be away from campus and using your own machine, or in an internet cafe to access them (a lot are blocked by universities through pressure from publishers). For some, like [Sci-hub](https://sci-hub.se/), you will need the [DOI](DOI.html) (you can get this from the publisher’s website). At the time of writing, Sci-hub has >85,483,812 papers. [Unpaywall.org](https://unpaywall.org/) has >28,112,995 papers. [Open Access Button](https://openaccessbutton.org/) don’t say how many they can reach, but “millions” of papers. And [freefullpdf](http://www.freefullpdf.com/) is a website that searches the internet and indicates where you can download pdfs (but sometimes you might hit a paywall).

@@

The chances are that you have friends at other universities where the subscription you are looking for is not behind a paywall. If not then ask your advisor or labmates, or post it on your lab’s (electronic) notice board.

### Go back to you librarian and ask them for help

>>>>

Once you understand where the money for science and publishing comes from, you will understand that there is no real shame in trying to access papers that are behind a paywall. Essentially, I’m going to recommend that you move from legitimate to illegitimate means, and then if all else fails to get an interlibrary loan. This is last on the list as someone will have to pay for it and it will take some time. Your library may have an allocation that they can use before the bill comes to your advisor’s desk.

@@

Google Scholar has lots of links (on the right) labelled with “[pdf]”, and these are often integrated with your library. If there’s nothing at first, click on the “All 9 versions” link (below and to the right). This may reveal that there is more than one version with a pdf, and you might have access to one of the others. Check for a link to the author on google scholar, and then look for a link to their website.

@@

### Write to the author and request a pdf
Some authors will respond very quickly to reprint requests, while you won't hear from others for sometime, or at all. However, this was the traditional way of obtaining a copy that you didn't have access to, and is definitely worth trying. You should be able to get the email address of the corresponding author from the journal website. Sometimes these are hidden and you need to jump through hoops to find them. You could also search for the email addresses of any co-authors, and also write to them.

@@

There are a few of these, and some have plugins to web browsers. You may need to be away from campus and using your own machine, or in an internet cafe to access them (a lot are blocked by universities through pressure from publishers). For some, like [Sci-hub](https://sci-hub.se/), you will need the [DOI](#DOI) (you can get this from the publisher’s website). At the time of writing, Sci-hub has >85,483,812 papers. [Unpaywall.org](https://unpaywall.org/) has >28,112,995 papers. [Library Genious](http://libgen.li/) or Libgen claim to have 3 million files. [Open Access Button](https://openaccessbutton.org/) don’t say how many they can reach, but “millions” of papers. And [freefullpdf](http://www.freefullpdf.com/) is a website that searches the internet and indicates where you can download pdfs (but sometimes you might hit a paywall).

@@

The chances are that you have friends at other universities where the subscription you are looking for is not behind a paywall. If not, then ask your advisor or labmates, or post it on your lab’s (electronic) notice board.

### Go back to your librarian and ask them for help
